\documentclass[letterpaper,12pt]{article}

\usepackage{threeparttable}
\usepackage{geometry}
\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1.25in,rmargin=1.25in}
\usepackage[format=hang,font=normalsize,labelfont=bf]{caption}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{array}
\usepackage{delarray}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{lscape}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{float,color}
\usepackage[pdftex]{graphicx}
\usepackage{pdfsync}
\usepackage{verbatim}
\usepackage{placeins}
\usepackage{geometry}
\usepackage{pdflscape}
\synctex=1
\usepackage{hyperref}
\hypersetup{colorlinks,linkcolor=red,urlcolor=blue,citecolor=red}
\usepackage{bm}


\theoremstyle{definition}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}{Definition} % Number definitions on their own
\newtheorem{derivation}{Derivation} % Number derivations on their own
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}{Proposition} % Number propositions on their own
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\bibliographystyle{aer}
\newcommand\ve{\epsilon}
%\renewcommand\theenumi{\roman{enumi}}
\newcommand\norm[1]{\left\lVert#1\right\rVert}

\begin{document}

\title{Algorithms for solving for SS and Time Path of OG model with heterogeneous static firms}
\date{\today}
\author{}
\maketitle

\begin{spacing}{1.5}
\pagenumbering{arabic}


The model assumed for these solutions is one without taxes or bequests.  Endogenous labor.  $J$ ability types.  $M$ firms.  $I$ consumption goods.  $T$ periods from initial state to the steady state.

\section{Steady State}

\begin{enumerate}
\item Guess $\bar{r}$ and $\bar{w}$
\item Determine output prices, $\bar{p}_{m}$ using zero profit condition and firm FOCS together. These imply:

\begin{equation}
\bar{p}_{m}  = \left[(1-\gamma_{m})\left(\frac{\bar{w}}{\bar{A}_{m}}\right)^{1-\epsilon_{m}} + \gamma_{m}\left(\frac{(\bar{r}+\delta_{m})}{\bar{A}_{m}}\right)^{1-\epsilon_{m}} \right]^{\frac{1}{1-\epsilon_{m}}}
\end{equation}
\item Determine consumption goods prices using the fixed coefficient matrix $\Pi$ mapping output goods to consumption goods: $\bar{p}^{c}_{i} = \sum_{m=1}^{M}\pi_{i,m}\bar{p}_{m}$
\item Determine the price of the composite consumption good: $\tilde{p} =  \prod_{i=1}^{I}\left( \frac{\bar{p}^{c}_{i}}{\alpha_{i}} \right)^{\alpha_{i}}$
\item With prices $\bar{r}, \bar{w}, \bar{p}^{c}_{i}, \tilde{p}$, we can solve the HH problem.  We can solve the problem for each type $j$ separately using a root-finding algorithm to solve for the $2\times S$ unknowns from $2\times S$ equations: the HH FOCs for the choices of savings and labor supply in each year of life.  This yields $\bar{b}_{j,s}$ and $\bar{n}_{j,s}$.
\item Using the HH's budget constraint and the vector of prices, we can solve for composite consumption, $\tilde{c}_{j,s}$.
\item We can solve for the total demand for each consumption good using the necessary conditions from the HH's subutility function.  These imply (where $\bar{c}_{i}$ is the minimum expenditure on good $i$): 

\begin{equation}
\bar{c}_{i,j,s}  = \frac{\alpha_{i} \tilde{p}_{s}\tilde{c}_{j,s}}{\bar{p}^{c}_{i}} + \bar{c}_{i},
\end{equation}

\item Summing over $J$ and $S$ we can get total demand for consumption of good $i$: $\bar{C}_{i} = \sum_{j=1}^{J}\sum_{s=1}^{S} \bar{c}_{i,j,s}$
\item With $\bar{C}_{i}$ and prices, we can solve for the demand for output from each industry using the resource constraint, the fixed coefficient matrices $\Pi$ and $\Xi$ (where $\Xi$ is the fixed coefficient input-output matrix - mapping production goods to capital goods), and the demand for investment.
	\begin{itemize}
	\item Investment demand for industry $m$ in the SS is given by: $\bar{I}_{m} = \delta_{m}\bar{K}_{m}$
	\item Demand for industry $m$ output from consumption is given by: $\bar{X}^{c}_{m} = \sum_{i=1}^{I} \pi_{i,m}\bar{C}_{i}$
	\item Demand for industry $m$ output from investment demand is given by: $\bar{X}^{i}_{m} =  \sum_{j=1}^{M} \xi_{j,m}\bar{I}_{j} = \sum_{j=1}^{M} \xi_{j,m}\delta_{j}\bar{K}_{j} $
	\item And we can use the firm's FOC for capital, labor, and the production function together,  to write the demand for capital as a function of output and factor prices: 
	
	\begin{equation}
	\label{eqn:k_demand2}
	\bar{K}_{m} = \frac{\bar{X}_{m}}{\bar{A}_{m}}\left[\gamma_{m}^{\frac{1}{\epsilon_{m}}}+(1-\gamma_{m})^{\frac{1}{\epsilon_{m}}}\left(\frac{\bar{r}+\delta_{m}}{\bar{w}}\right)^{\epsilon_{m}-1}\left(\frac{1-\gamma_{m}}{\gamma_{m}}	\right)^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}\right]^{\frac{\epsilon_{m}}{1-\epsilon_{m}}}
	\end{equation}
	
	\item Resource constraint for industry $m$: 
	\begin{equation}
	\begin{split}
	\bar{X}_{m} &= \bar{X}^{c}_{m} + \bar{X}^{i}_{m} \\
	&= \sum_{i=1}^{I} \pi_{i,m}\bar{C}_{i} +\sum_{j=1}^{M} \xi_{j,m}\bar{I}_{j} \\
	 &= \sum_{i=1}^{I} \pi_{i,m}\bar{C}_{i} +\sum_{j=1}^{M} \xi_{j,m}\delta_{j}\bar{K}_{j}  \\
	 &= \sum_{i=1}^{I} \pi_{i,m}\bar{C}_{i} +\sum_{j=1}^{M} \xi_{j,m}\delta_{j} \frac{\bar{X}_{j}}{\bar{A}_{j}}\left[\gamma_{j}^{\frac{1}{\epsilon_{j}}}+(1-\gamma_{j})^{\frac{1}{\epsilon_{j}}}\left(\frac{\bar{r}+\delta_{j}}{\bar{w}}\right)^{\epsilon_{j}-1}\left(\frac{1-\gamma_{j}}{\gamma_{j}}	\right)^{\frac{\epsilon_{j}-1}{\epsilon_{j}}}\right]^{\frac{\epsilon_{j}}{1-\epsilon_{j}}}
	\end{split}
	\end{equation}
	
	\item The above is a system of $M$ equations and $M$ unknowns: a root finder can be used to solve for $\bar{X}_{m}$
	\end{itemize}

\item With $\bar{X}_{m}$ and prices, we can solve for the demand for capital from each industry.  We'll use the capital market clearing condition ($\sum_{m=1}^{M} \bar{K}_{m}=\sum_{j=1}^{J}\sum_{s=1}^{S} \bar{b}_{j,s}$) and the firm's FOC for capital together to find these demands.  Denoting the supply of capital as $\bar{K}^{s}=\sum_{j=1}^{J}\sum_{s=1}^{S} \bar{b}_{j,s}$, we have:

	\begin{equation}
	\bar{K}_{m} = \bar{K}^{s} - \sum_{k \neq m} \gamma_{k}\bar{X}_{k}\left(\frac{\bar{p}_{m}\left(\frac{\gamma_{m}\bar{X}_{m}}{\bar{K}_{m}}\right)\bar{A}_{m}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}+\delta_{k}}{\bar{p}_{k}}\bar{A}_{k}^{\frac{1-\epsilon_{k}}{\epsilon_{k}}}\right)^{-\epsilon_{k}}
	\end{equation}
	This is a system of $M$ nonlinear equations that can be solved for $\bar{K}_{m}$ using a root finding algorithm.
	
\item With $\bar{X}_{m}$ and prices, we can solve for the demand for labor from each industry.  We'll use the labor market clearing condition ($\sum_{m=1}^{M} \bar{EL}_{m}=\sum_{j=1}^{J}\sum_{s=1}^{S}e_{j,s} \bar{n}_{j,s}$) and the firm's FOC for labor together to find these demands.  Denoting the supply of labor as $\bar{L}^{s}=\sum_{j=1}^{J}\sum_{s=1}^{S} e_{j,s}\bar{n}_{j,s}$, we have:

	\begin{equation}
	\bar{EL}_{m} = \bar{L}^{s} - \sum_{k\neq m} (1- \gamma_{k})\bar{X}_{k}\left(\frac{\bar{p}_{m}\left(\frac{(1-\gamma_{m})\bar{X}_{m}}{\bar{EL}_{m}}\right)\bar{A}_{m}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}+\delta_{k}}{\bar{p}_{k}}\bar{A}_{k}^{\frac{1-\epsilon_{k}}{\epsilon_{k}}}\right)^{-\epsilon_{k}}
	\end{equation}
	This is a system of $M$ nonlinear equations that can be solved for $\bar{EL}_{m}$ using a root finding algorithm.
	
\item Use the demands for capital and labor found above in the FOC for a particular firm $m$ (can be any firm), this will imply and interest rate and wage rate.  Call these $r_{new}$ and $w_{new}$, respectively.  We have:

	\begin{equation}
	r_{new} = \bar{p}_{m}\left(\frac{\gamma_{m}\bar{X}_{m}}{\bar{K}_{m}}\right)\bar{A}_{m}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}, \forall \ m
	\end{equation}

	\begin{equation}
	w_{new} = \bar{p}_{m}\left(\frac{(1-\gamma_{m})\bar{X}_{m}}{\bar{EL}_{m}}\right)\bar{A}_{m}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}, \forall \ m
	\end{equation}

\item If $r_{new}$ and $w_{new}$ are equal to $\bar{r}$ and $\bar{w}$ guessed, then stop.  Else, update the guesses over $\bar{r}$ and $\bar{w}$ and repeat.  One can use $r_{new}$ and $w_{new}$ to inform the new guess.

\end{enumerate}



\section{Time Path}

\begin{enumerate}
\item Guess the time paths for factor prices: $r_{t}$ and $w_{t}$
\item Determine time path for output prices, $p_{m,t}$ using zero profit condition and firm FOCS together. These imply:

\begin{equation}
p_{m,t}  = \left[(1-\gamma_{m})\left(\frac{w_{t}}{A_{m,t}}\right)^{1-\epsilon_{m}} + \gamma_{m}\left(\frac{(r_{t}+\delta_{m})}{A_{m,t}}\right)^{1-\epsilon_{m}} \right]^{\frac{1}{1-\epsilon_{m}}}
\end{equation}
\item Determine the time path for consumption goods prices using the fixed coefficient matrix $\Pi$ mapping output goods to consumption goods: $p^{c}_{i,t} = \sum_{m=1}^{M}\pi_{i,m}p_{m,t}$
\item Determine time path for the price of the composite consumption good: $\tilde{p}_{t} =  \prod_{i=1}^{I}\left( \frac{p^{c}_{i,t}}{\alpha_{i}} \right)^{\alpha_{i}}$
\item With the time paths of prices $r_{t}, w_{t}, p^{c}_{i,t}, \tilde{p}_{t}$, we can solve the HH problem.  I think we can solve the path for each type $j$ and cohort separately using a root-finding algorithm to solve for the $2\times S$ unknowns from $2\times S$ equations: the HH FOCs for the choices of savings and labor supply.  This yields, $b_{j,s,t}$ and $n_{j,s,t}$.
\item Using the HH's budget constraint and the vector of prices, we can solve for composite consumption, $\tilde{c}_{j,s,t}$.
\item We can solve for the total demand for each consumption good using the necessary conditions from the HH's subutility function.  These imply (where $\bar{c}_{i}$ is the minimum expenditure on good $i$): 

\begin{equation}
c_{i,j,s,t}  = \frac{\alpha_{i} \tilde{p}_{s,t}\tilde{c}_{j,s,t}}{p_{i,t}} + \bar{c}_{i},
\end{equation}

\item Summing over $J$ and $S$ we can get time path for the demand for consumption of good $i$: $C_{i,t} = \sum_{j=1}^{J}\sum_{s=1}^{S} c_{i,j,s,t}$
\item With $C_{i,t}$ and prices, we can solve for the demand for output from each industry using the resource constraint, the fixed coefficient matrices $\Pi$ and $\Xi$ (where $\Xi$ is the fixed coefficient input-output matrix - mapping production goods to capital goods), and the demand for investment.
	\begin{itemize}
	\item Investment demand for industry $m$ in period $t$ is given by: $I_{m,t} = K_{m,t+1} - (1 - \delta_{m})K_{m,t}$
	\item Demand for industry $m$ output in period $t$ from consumption is given by: $X^{c}_{m,t} = \sum_{i=1}^{I} \pi_{i,m}C_{i,t}$
	\item Demand for industry $m$ output in period $t$ from investment demand is given by: $X^{i}_{m,t} =  \sum_{j=1}^{M} \xi_{j,m}I_{j,t} = \sum_{j=1}^{M} \xi_{j,m}\left(K_{j,t+1} - (1\delta_{m})K_{j,t}\right) $
	\item And we can use the firm's FOC for capital, labor, and the production function together,  to write the demand for capital in each period as a function of output and factor prices: 
	
	\begin{equation}
	\label{eqn:k_demand2}
	K_{m,t} = \frac{X_{m,t}}{A_{m,t}}\left[\gamma_{m}^{\frac{1}{\epsilon_{m}}}+(1-\gamma_{m})^{\frac{1}{\epsilon_{m}}}\left(\frac{r_{t}+\delta_{m}}{w_{t}}\right)^{\epsilon_{m}-1}\left(\frac{1-\gamma_{m}}{\gamma_{m}}	\right)^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}\right]^{\frac{\epsilon_{m}}{1-\epsilon_{m}}}
	\end{equation}
	
	\item Resource constraint for industry $m$ in year $t$: 
	\begin{equation}
	\begin{split}
	X_{m,t} &= X^{c}_{m,t} + X^{i}_{m,t} \\
	&= \sum_{i=1}^{I} \pi_{i,m}C_{i,t} +\sum_{j=1}^{M} \xi_{j,m}I_{j,t} \\
	 &= \sum_{i=1}^{I} \pi_{i,m}C_{i,t} +\sum_{j=1}^{M} \xi_{j,m}\left(K_{j,t+1} - (1\delta_{j})K_{j,t} \right)  \\
	 &= \sum_{i=1}^{I} \pi_{i,m}C_{i,t} ... \\
	 & \quad +\sum_{j=1}^{M} \xi_{j,m}\left[\left( \frac{X_{j,t+1}}{A_{j,t+1}}\left[\gamma_{j}^{\frac{1}{\epsilon_{j}}}+(1-\gamma_{j})^{\frac{1}{\epsilon_{j}}}\left(\frac{r_{t+1}+\delta_{j}}{w_{t+1}}\right)^{\epsilon_{j}-1}\left(\frac{1-\gamma_{j}}{\gamma_{j}}\right)^{\frac{\epsilon_{j}-1}{\epsilon_{j}}}\right]^{\frac{\epsilon_{j}}{1-\epsilon_{j}}}\right) ... \right. \\
	 & \left. \quad - (1-\delta_{j})\left( \frac{X_{j,t}}{A_{j,t}}\left[\gamma_{j}^{\frac{1}{\epsilon_{j}}}+(1-\gamma_{j})^{\frac{1}{\epsilon_{j}}}\left(\frac{r_{t}+\delta_{j}}{w_{t}}\right)^{\epsilon_{j}-1}\left(\frac{1-\gamma_{j}}{\gamma_{j}}	\right)^{\frac{\epsilon_{j}-1}{\epsilon_{j}}}\right]^{\frac{\epsilon_{j}}{1-\epsilon_{j}}}\right) \right]
	\end{split}
	\end{equation}
	
	\item The above is a system of $M\times T$ equations and $M\times T$ unknowns: a root finder can be used to solve for $X_{m,t}$
	\item Alternatively, one can leverage the recursive nature of the problem.  In particular, since we know the steady state solution, we have $K_{m,T}$ for all $m$.  Thus we can find $I_{m,T-1} = K_{m,T} - (1\delta_{m})K_{m,T-1}$.  Then using the equations above, we can solve a system of $M$ equations and $M$ unknowns to find $X_{m,T-1}$.  We then repeat this for each period from $T-1$ back to the initial period.  This yields the time path of output demand, $X_{m,t}$ without having to solve a very large system of equations at once.
	\end{itemize}

\item With the time path for output demand, $X_{m,t}$, and prices, we can solve for the demand for capital from each industry in each period $t$.  We'll use the capital market clearing condition ($\sum_{m=1}^{M} K_{m,t}=\sum_{j=1}^{J}\sum_{s=1}^{S} b_{j,s,t}$) and the firm's FOC for capital together to find these demands.  Denoting the supply of capital as $K^{s}_{t}=\sum_{j=1}^{J}\sum_{s=1}^{S} b_{j,s,t}$, we have:

	\begin{equation}
	K_{m,t} = K^{s,t} - \sum_{k \neq m} \gamma_{k}X_{k,t}\left(\frac{p_{m,t}\left(\frac{\gamma_{m}X_{m,t}}{K_{m,t}}\right)A_{m,t}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}+\delta_{k}}{p_{k,t}}A_{k,t}^{\frac{1-\epsilon_{k}}{\epsilon_{k}}}\right)^{-\epsilon_{k}}
	\end{equation}
	For each year, $t$, this is a system of $M$ nonlinear equations that can be solved for $K_{m,t}$ using a root finding algorithm.
	
\item With the time path for output demand, $X_{m,t}$ and prices, we can solve for the demand for labor from each industry.  We'll use the labor market clearing condition ($\sum_{m=1}^{M} EL_{m,t}=\sum_{j=1}^{J}\sum_{s=1}^{S}e_{j,s} n_{j,s,t}$) and the firm's FOC for labor together to find these demands.  Denoting the supply of labor as $L^{s,t}=\sum_{j=1}^{J}\sum_{s=1}^{S} e_{j,s}n_{j,s,t}$, we have:

	\begin{equation}
	EL_{m,t} = L^{s,t} - \sum_{k\neq m} (1- \gamma_{k})X_{k,t}\left(\frac{p_{m,t}\left(\frac{(1-\gamma_{m})X_{m,t}}{EL_{m,t}}\right)A_{m,t}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}+\delta_{k}}{p_{k,t}}A_{k,t}^{\frac{1-\epsilon_{k}}{\epsilon_{k}}}\right)^{-\epsilon_{k}}
	\end{equation}
	For each year, $t$, this is a system of $M$ nonlinear equations that can be solved for $EL_{m,t}$ using a root finding algorithm.
	
\item Use the demands for capital and labor found above in the FOC for a particular firm, this will imply and interest rate and wage rate in each period, $t$.  Call these $r_{new,t}$ and $w_{new,t}$, respectively.  We have:

	\begin{equation}
	r_{new,t} = p_{m}\left(\frac{\gamma_{m}X_{m,t}}{K_{m,t}}\right)A_{m,t}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}-\delta_{m}, \forall \ m, t
	\end{equation}

	\begin{equation}
	w_{new} = p_{m,t}\left(\frac{(1-\gamma_{m})X_{m,t}}{EL_{m,t}}\right)A_{m,t}^{\frac{\epsilon_{m}-1}{\epsilon_{m}}}, \forall \ m, t
	\end{equation}

\item If $r_{new,t}$ and $w_{new,t}$ are equal the time paths $r_{t}$ and $w_{t}$ guessed, then stop.  Else, update the guesses over $r_{t}$ and $w_{t}$ and repeat.  One can use $r_{new,t}$ and $w_{new,t}$ to inform the new guess. 

\end{enumerate}

\end{spacing}
\end{document}